---
title: "Kelleher Rhizo Alpha Diversity"
output: html_document
date: "2024-05-10"
---

Load Packages
```{r}
library(vegan)
library(ggplot2)
library(tidyverse)
library(readxl)
library(RColorBrewer)

set.seed(13)
```

# Alpha Diversity Analysis
```{r}
ASV <- read.csv("~/ASVtable_rhizosbulk_forR_noCorM.csv", header=TRUE, row.names=1)
```

## Species richness
```{r}
ASV_rare_rhizobulk <- rarefy(ASV,4000)
#Create a new ASV file on your computer. You may want to write this link different depending on where you want the file to end up
write.csv(ASV_rare_rhizobulk,"~/ASV_speciesrichness_rhizobulk.csv", row.names=TRUE)
#Make a new file using the one just created to have the correct format. 
#Columns: Sample, Richness, metadata info and save to same location with a new name (add 1 to the end)
```

## Shannon diversity
```{r}
shannon_iteration <- function(){
    ASV %>%
    rrarefy(sample=2859) %>%
    diversity()
  
}
ASV_shannon_rhizobulk <- replicate(100, shannon_iteration())
#Create a new ASV file on your computer. You may want to write this link different depending on where you want the file to end up
write.csv(ASV_shannon_rhizobulk,"~/ASV_shannon_rhizobulk.csv", row.names=TRUE)
#Make a new file using the one just created to have the correct format. 
#Columns: Sample, Shannon, metadata info and save to same location with a new name (add 1 to the end)
```

Now test the significance between treatments using the Kuksal-Wallis test since the data is non-parametric
```{r}
species_rich <- read.csv("~/ASV_speciesrichness_rhizobulk1.csv")
shannon <- read.csv("~/ASV_shannon_rhizobulk1.csv")
```

## Species richness
### Soil type
```{r}
kruskal.test(Richness ~ Depth, data=species_rich)
```
### Grassland
```{r}
kruskal.test(Richness ~ Grassland, data=species_rich)
```

### Month
```{r}
kruskal.test(Richness ~ Date, data=species_rich)
```
### Site
```{r}
kruskal.test(Richness ~ Site, data=species_rich)
```

### Year
```{r}
kruskal.test(Richness ~ Year, data=species_rich)
```
## Shannon Diversity
### Soil Type
```{r}
kruskal.test(Shannon ~ Depth, data=shannon)
```
### Grassland
```{r}
kruskal.test(Shannon ~ Grassland, data=shannon)
```

### Month
```{r}
kruskal.test(Shannon ~ Date, data=shannon)
```
### Site 
```{r}
kruskal.test(Shannon ~ Site, data=shannon)
```

### Year
```{r}
kruskal.test(Shannon ~ Year, data=shannon)
```
# Box and Jitter Plots
Load packages
```{r}
library(ggplot2)
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(glue)
library(ggtext)

set.seed(13)
```

## Soil Type
### Species Richness
```{r}
ggplot(species_rich, aes(x=reorder(Depth,Richness), y=Richness, fill=Depth)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Species Richness",x=NULL, y="Species Richness") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("Bulk","Proximal","Rhizosphere"),
                    labels=c("Bulk","Proximal","Rhizosphere"),
                    values=c("pink", "blue", "darkgreen")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
### Shannon Diversity
```{r}
ggplot(shannon, aes(x=reorder(Depth,Shannon), y=Shannon, fill=Depth)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Shannon Index",x=NULL, y="Shannon Diversity Index") + 
  scale_fill_manual(name=NULL, 
                     breaks=c("Bulk","Proximal","Rhizosphere"),
                    labels=c("Bulk","Proximal","Rhizosphere"),
                    values=c("pink", "blue", "darkgreen")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```

## Grassland
### Species Richness
```{r}
ggplot(species_rich, aes(x=reorder(Grassland,Richness), y=Richness, fill=Grassland)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Species Richness",x=NULL, y="Species Richness") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("Nesco","Ricketts"),
                    labels=c("Nescopeck","Ricketts"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
### Shannon Diversity
```{r}
ggplot(shannon, aes(x=reorder(Grassland,Shannon), y=Shannon, fill=Grassland)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Shannon Index",x=NULL, y="Shannon Diversity Index") + 
  scale_fill_manual(name=NULL, 
                     breaks=c("Nesco","Ricketts"),
                    labels=c("Nescopeck","Ricketts"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```

## Month
### Species Richness
```{r}
ggplot(species_rich, aes(x=reorder(Date,Richness), y=Richness, fill=Date)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Species Richness",x=NULL, y="Species Richness") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("June","August"),
                    labels=c("June","August"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
### Shannon Diversity
```{r}
ggplot(shannon, aes(x=reorder(Date,Shannon), y=Shannon, fill=Date)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Shannon Index",x=NULL, y="Shannon Diversity Index") + 
  scale_fill_manual(name=NULL, 
                     breaks=c("June","August"),
                    labels=c("June","August"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```


## Site
### Species Richness
```{r}
ggplot(species_rich, aes(x=reorder(Site,Richness), y=Richness, fill=Site)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Species Richness",x=NULL, y="Species Richness") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("Site 1","Site 2","Site 3"),
                    labels=c("Ricketts Site 1","Ricketts Site 2","Nescopeck"),
                    values=c("pink", "blue", "darkgreen")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
### Shannon Diversity
```{r}
ggplot(shannon, aes(x=reorder(Site,Shannon), y=Shannon, fill=Site)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Shannon Index",x=NULL, y="Shannon Diversity Index") + 
  scale_fill_manual(name=NULL, 
                     breaks=c("Site 1","Site 2","Site 3"),
                    labels=c("Ricketts Site 1","Ricketts Site 2","Nescopeck"),
                    values=c("pink", "blue", "darkgreen")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```

## Year
### Species Richness
```{r}
ggplot(species_rich, aes(x=reorder(Year,Richness), y=Richness, fill=Year)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Species Richness",x=NULL, y="Species Richness") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("Year1","Year2"),
                    labels=c("2021", "2022"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
### Shannon Diversity
```{r}
ggplot(shannon, aes(x=reorder(Year,Shannon), y=Shannon, fill=Year)) + 
  geom_boxplot(show.legend = FALSE, outlier.shape=NA, alpha=0.25, width=0.6, coef=0) +
  geom_jitter(show.legend = FALSE, width = 0.25, shape=21, size=2, color="black") +
  labs(title="Shannon Index",x=NULL, y="Shannon Diversity Index") + 
  scale_fill_manual(name=NULL, 
                    breaks=c("Year1","Year2"),
                    labels=c("2021", "2022"),
                    values=c("pink", "blue")) +
  theme_classic()+theme(plot.title = element_text(hjust=0.5))
```
