---
title: "Lily_Rhizosbulk_Indicator"
output: html_document
date: "2024-05-14"
---
Load Packages
```{r}
library(permute)
library(lattice)
library(vegan)
library(ggplot2)
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(indicspecies)

set.seed(13)
```

```{r}
ASV <- read.csv("~/ASVtable_rhizosbulk_forR_noCorM.csv", header=TRUE)

metadata <- read.csv("~/Lily_Rhizo_Metadata_modified.csv")

taxonomy <- read.csv("~/ASV_taxonomy_rhizosbulk_noCorM.csv") %>% rename("ASV" =X)
```

```{r}

ASV_meta_relabund <- inner_join(metadata, ASV, by="Sample")

abund <- ASV_meta_relabund[,8:ncol(ASV_meta_relabund)]
grass <- ASV_meta_relabund$Grassland
site <- ASV_meta_relabund$Site
soiltype <- ASV_meta_relabund$Depth
date <- ASV_meta_relabund$Date
year <- ASV_meta_relabund$Year
ph <- ASV_meta_relabund$ph
```

```{r}
indspecies_byph <- multipatt(abund, ph, control = how(nperm=999),duleg = TRUE)
summary(indspecies_byph, alpha=0.01, minstat=0.8)
```
Low ph:
ASV 63, 125, 56

high ph:
ASV 1631, 2519, 228



```{r}
indspecies_bygrass <- multipatt(abund, grass, control = how(nperm=999),duleg = TRUE)
summary(indspecies_bygrass, alpha=0.01, minstat=0.8)
```
Indicators by grassland: 

Nesco: 24, 31, 30, 2

Ricketts: 16, 5, 10, 82

```{r}
indspecies_bysite <- multipatt(abund, site, control = how(nperm=999),duleg = TRUE)
summary(indspecies_bysite, alpha=0.01, minstat=0.8)
```

Indicators by site: 

Site 1: 27, 76, 84, 77

Site 2: 80, 90, 324, 99

Nescopeck: 24, 30, 31, 352

```{r}
indspecies_bysoiltype <- multipatt(abund, soiltype, control = how(nperm=999),duleg = TRUE)
summary(indspecies_bysoiltype, alpha=0.01, minstat=0.8)
```

Indicators by soil type: 

Rhizosphere: 303, 110

```{r}
indspecies_bymonth <- multipatt(abund, date, control = how(nperm=999),duleg = TRUE)
summary(indspecies_bymonth, alpha=0.01, minstat=0.8)
```

```{r}
indspecies_byyear <- multipatt(abund, year, control = how(nperm=999),duleg = TRUE)
summary(indspecies_byyear, alpha=0.01, minstat=0.8)
```

Indicators by year: 

2021: 5, 16, 10, 13

## By Just grassland look at soil typefro just ricketts and just nescopeck


# Preparing bubble plot
```{r}
ASV_B <- read.csv("~/ASVtable_rhizosbulk_forR_noCorM.csv", header=TRUE, row.names=1)
dist_matrix <- avgdist(ASV_B,sample=4000, dmethod = "bray")
soil_nmds <- metaMDS(dist_matrix) %>% scores() %>% as_tibble(rownames = "Sample")

```

```{r}
bubble <- inner_join(soil_nmds, ASV_meta_relabund, by="Sample")
```

## Ph BubblePlot
Low ph:
ASV 63, 125, 56

high ph:
ASV 1631, 2519, 228

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_63, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV63, Phylum=*Acidobacteriota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_125, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV125, Phylum=*Planctomycetota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```


```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_56, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV56, Phylum=*Gammaproteobacteria* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_1631, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV1631, Phylum=*Myxococcota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_2519, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV2519, Phylum=*Bacteroidota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_228, fill=ph)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("red","blue", "green"), 
                      breaks = c("Four", "Five", "Six"),
                      labels = c("pH 4-5", "pH 5-6", "pH 6-7")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV228, Phylum=*Latescibacterota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

## Grassland Bubbleplot
Nesco: 24, 31, 30, 2

Ricketts: 16, 5, 10, 82
```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_24, fill=Grassland)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("pink","blue"), 
                      breaks = c("Nesco", "Ricketts"),
                      labels = c("Nescopeck", "Ricketts")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV24, Phylum=*Bacteroidetes* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```
## Soil Type Bubble Plot
```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_303, fill=Depth)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("violet","blue","orange"), 
                      breaks = c("Bulk", "Proximal", "Rhizosphere"),
                      labels = c("Bulk", "Proximal", "Rhizosphere")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV303, Phylum=*Proteobacteria* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```

```{r}
ggplot(bubble, aes(x=NMDS1, y=NMDS2, size=ASV_110, fill=Depth)) + 
    geom_point(alpha=0.7, shape=21, color="black", stroke = 0.75) +  
    scale_fill_manual(name=NULL, values = c("violet","blue","orange"), 
                      breaks = c("Bulk", "Proximal", "Rhizosphere"),
                      labels = c("Bulk", "Proximal", "Rhizosphere")) + 
    guides(fill = guide_legend(override.aes = list(size = 6, alpha=1))) + 
    scale_size(range = c(1, 15), name="Relative Abundance (%)") +
    labs(title = "Relative abundance of ASV110, Phylum=*Myxococcota* ", x = "Dimension 1", y = "Dimension 2", 
         subtitle = "Experimental Indicactor ASV") +
    theme_classic() + 
    theme(legend.text= element_text(size=10),
        legend.key.size = unit(15, "pt"))
```


