---
title: "Kelleher Rhizo Permanova"
output: html_document
date: "2024-05-09"
---
Load Packages
```{r}
library(mgcv)
library(nlme)
library(permute)
library(lattice)
library(vegan)
library(ggplot2)
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(ecole)


set.seed(13)
```

Load in Data
```{r}
ASV <- read.csv("~/ASVtable_rhizosbulk_forR_noCorM.csv", header=TRUE, row.names=1)
metadata <- read.csv("~/Lily_Rhizo_Metadata_modified.csv")
```

Create a dissimilarity matrix
```{r}
rhizosbulk_dis <- avgdist(ASV,sample=4000, dmethod = "bray")
```

## Run Permanovas
### Grassland
```{r}
perm_grassland <-adonis2(rhizosbulk_dis~as.factor(metadata$Grassland), data=metadata, permutations=9999)
perm_grassland #Viewing data
```
### Soil Type
```{r}
perm_soiltype <-adonis2(rhizosbulk_dis~as.factor(metadata$Depth), data=metadata, permutations=9999)
perm_soiltype #Viewing data
```
Pairwise!
```{r}
soiltype_pvalue_table <- permanova_pairwise(rhizosbulk_dis, metadata$Depth, permutations=9999 ,padj = "bonferroni")
soiltype_pvalue_table
```

### Month
```{r}
perm_month <-adonis2(rhizosbulk_dis~as.factor(metadata$Date), data=metadata, permutations=9999)
perm_month #Viewing data
```

### Year
```{r}
perm_year <-adonis2(rhizosbulk_dis~as.factor(metadata$Year), data=metadata, permutations=9999)
perm_year #Viewing data
```
### Site
```{r}
perm_site <-adonis2(rhizosbulk_dis~as.factor(metadata$Site), data=metadata, permutations=9999)
perm_site #Viewing data
```
Pairwise!
```{r}
site_pvalue_table <- permanova_pairwise(rhizosbulk_dis, metadata$Site, permutations=9999 ,padj = "bonferroni")
site_pvalue_table
```
### pH
```{r}
perm_ph <-adonis2(rhizosbulk_dis~as.factor(metadata$ph), data=metadata, permutations=9999)
perm_ph #Viewing data
```
```{r}
ph_pvalue_table <- permanova_pairwise(rhizosbulk_dis, metadata$ph, permutations=9999 ,padj = "bonferroni")
ph_pvalue_table
```

## NMDS
```{r}
nmds <- metaMDS(rhizosbulk_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds)

stress_value <- metaMDS(rhizosbulk_dis)$stress
stress_value

```
Add in metadata
```{r}
metadata_nmds <- inner_join(metadata, nmds)
plot(metadata_nmds)
```

### Grassland
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Grassland)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkgreen", "blue"), 
                     breaks = c("Nesco", "Ricketts"), 
                     labels = c("Nescopeck", "Ricketts")) +
  labs(title = "Microbial Communities Separated by Grassland", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

### PH
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=ph)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("red", "blue"), 
                     breaks = c("Four", "Five"), 
                     labels = c("pH below 5", "pH 5 and above")) +
  labs(title = "Microbial Communities Separated by pH", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```


### Soil Type
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Depth)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("violet", "blue", "orange"), 
                     breaks = c("Bulk", "Rhizosphere", "Proximal"), 
                     labels = c("Bulk", "Rhizosphere", "Proximal")) +
  labs(title = "Microbial Communities Separated by Soil type", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
ggsave("All_Samples_SoilType.png", height=4, width=5, device="png") 
```

### Month
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Date)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkblue", "red"), 
                     breaks = c("August", "June"), 
                     labels = c("August", "June")) +
  labs(title = "Microbial Communities Separated by Month", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

### Year
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Year)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("red", "blue"), 
                     breaks = c("Year1", "Year2"), 
                     labels = c("2021", "2022")) +
  labs(title = "Microbial Communities Separated by Year", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```
### Site
```{r}
metadata_nmds %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Site)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkgreen", "blue", "purple"), 
                     breaks = c("Site 1", "Site 2", "Site 3"), 
                     labels = c("Ricketts Site 1", "Ricketts Ste 2", "Nescopeck")) +
  labs(title = "Microbial Communities Separated by Grassland Site", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

## Does Grassland matter by soil type??
### Bulk
Load in Data
```{r}
ASV_bulk <- read.csv("~/ASVtable_rhizosbulk_forR_bulk.csv", header=TRUE, row.names=1)
metadata_bulk <- read.csv("~/Lily_Rhizo_Metadata_bulk.csv")
```

Create a dissimilarity matrix
```{r}
bulk_dis <- avgdist(ASV_bulk,sample=4110, dmethod = "bray")
```

Grassland
```{r}
perm_soiltype_bulk <-adonis2(bulk_dis~as.factor(metadata_bulk$Grassland), data=metadata_bulk, permutations=9999)
perm_soiltype_bulk #Viewing data
```
NMDS
```{r}
nmds_bulk <- metaMDS(bulk_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds_bulk)
```
Add in metadata
```{r}
metadata_nmds_bulk <- inner_join(metadata_bulk, nmds_bulk)
plot(metadata_nmds_bulk)
```
```{r}
metadata_nmds_bulk %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Grassland)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkgreen", "blue"), 
                     breaks = c("Nesco", "Ricketts"), 
                     labels = c("Nescopeck", "Ricketts")) +
  labs(title = "Bulk Soil Microbial Communities Separated by Grassland", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

### Proximal
Load in Data
```{r}
ASV_proximal <- read.csv("~/ASVtable_rhizosbulk_forR_proximal.csv", header=TRUE, row.names=1)
metadata_proximal <- read.csv("~/Lily_Rhizo_Metadata_proximal.csv")
```

Create a dissimilarity matrix
```{r}
proximal_dis <- avgdist(ASV_proximal,sample=4400, dmethod = "bray")
```

Grassland
```{r}
perm_soiltype_proximal <-adonis2(proximal_dis~as.factor(metadata_proximal$Grassland), data=metadata_proximal, permutations=9999)
perm_soiltype_proximal #Viewing data
```
NMDS
```{r}
nmds_proximal <- metaMDS(proximal_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds_proximal)
```
Add in metadata
```{r}
metadata_nmds_proximal <- inner_join(metadata_proximal, nmds_proximal)
plot(metadata_nmds_proximal)
```
```{r}
metadata_nmds_proximal %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Grassland)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkgreen", "blue"), 
                     breaks = c("Nesco", "Ricketts"), 
                     labels = c("Nescopeck", "Ricketts")) +
  labs(title = "Proximal Soil Microbial Communities Separated by Grassland", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

### Rhizosphere
Load in Data
```{r}
ASV_rhizo <- read.csv("~/ASVtable_rhizosbulk_forR_rhizo.csv", header=TRUE, row.names=1)
metadata_rhizo <- read.csv("~/Lily_Rhizo_Metadata_rhizo.csv")
```

Create a dissimilarity matrix
```{r}
rhizo_dis <- avgdist(ASV_rhizo,sample=4200, dmethod = "bray")
```

Grassland
```{r}
perm_soiltype_rhizo <-adonis2(rhizo_dis~as.factor(metadata_rhizo$Grassland), data=metadata_rhizo, permutations=9999)
perm_soiltype_rhizo #Viewing data
```
NMDS
```{r}
nmds_rhizo <- metaMDS(rhizo_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds_rhizo)
```
Add in metadata
```{r}
metadata_nmds_rhizo <- inner_join(metadata_rhizo, nmds_rhizo)
```
```{r}
metadata_nmds_bulk %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Grassland)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("darkgreen", "blue"), 
                     breaks = c("Nesco", "Ricketts"), 
                     labels = c("Nescopeck", "Ricketts")) +
  labs(title = "Rhizosphere Soil Microbial Communities Separated by Grassland", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

## Does soil type matter for each grasland??
### Ricketts
Load in Data
```{r}
ASV_ricketts <- read.csv("~/ASVtable_rhizosbulk_forR_Ricketts.csv", header=TRUE, row.names=1)
metadata_ricketts <- read.csv("~/Lily_Rhizo_Metadata_Ricketts.csv")
```

Create a dissimilarity matrix
```{r}
ricketts_dis <- avgdist(ASV_ricketts,sample=4110, dmethod = "bray")
```

Soil Type
```{r}
perm_soiltype_ricketts <-adonis2(ricketts_dis~as.factor(metadata_ricketts$Depth), data=metadata_ricketts, permutations=9999)
perm_soiltype_ricketts #Viewing data
```
Pairwise!
```{r}
soiltypericketts_pvalue_table <- permanova_pairwise(ricketts_dis, metadata_ricketts$Depth, permutations=9999 ,padj = "bonferroni")
soiltypericketts_pvalue_table
```


NMDS
```{r}
nmds_ricketts <- metaMDS(ricketts_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds_ricketts)

stress_value_rick <- metaMDS(ricketts_dis)$stress
stress_value_rick
```
Add in metadata
```{r}
metadata_nmds_ricketts <- inner_join(metadata_ricketts, nmds_ricketts)
plot(metadata_nmds_ricketts)
```

```{r}
metadata_nmds_ricketts %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Depth)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("violet", "blue", "orange"), 
                     breaks = c("Bulk", "Proximal", "Rhizosphere"), 
                     labels = c("Bulk", "Proximal", "Rhizosphere")) +
  labs(title = "Microbial Communities In Ricketts Separated by Soil type", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```

### Nesco
Load in Data
```{r}
ASV_Nesco <- read.csv("~/ASVtable_rhizosbulk_forR_Nesco.csv", header=TRUE, row.names=1)
metadata_Nesco <- read.csv("~/Lily_Rhizo_Metadata_Nesco.csv")
```

Create a dissimilarity matrix
```{r}
nesco_dis <- avgdist(ASV_Nesco,sample=5730, dmethod = "bray")
```

Soil Type
```{r}
perm_soiltype_nesco <-adonis2(nesco_dis~as.factor(metadata_Nesco$Depth), data=metadata_Nesco, permutations=9999)
perm_soiltype_nesco #Viewing data
```
Pairwise!
```{r}
soiltypenesco_pvalue_table <- permanova_pairwise(nesco_dis, metadata_Nesco$Depth, permutations=9999 ,padj = "bonferroni")
soiltypenesco_pvalue_table
```


NMDS
```{r}
nmds_nesco <- metaMDS(nesco_dis) %>% scores() %>% as_tibble(rownames = "Sample")
plot(nmds_nesco)


stress_value_nesco <- metaMDS(nesco_dis)$stress
stress_value_nesco
```
Add in metadata
```{r}
metadata_nmds_nesco <- inner_join(metadata_Nesco, nmds_nesco)
plot(metadata_nmds_nesco)
```

```{r}
metadata_nmds_nesco %>% ggplot(aes(x=NMDS1,y=NMDS2, color=Depth)) + geom_point(shape = 19, size=3) + 
  coord_fixed() + 
  scale_color_manual(name=NULL, 
                     values = c("violet", "blue", "orange"), 
                     breaks = c("Bulk", "Proximal", "Rhizosphere"), 
                     labels = c("Bulk", "Proximal", "Rhizosphere")) +
  labs(title = "Microbial Communities In Nescopeck Separated by Soil type", x = "Dimension 1", y = "Dimension 2") + 
  theme_classic() + theme(plot.title = element_text(size=15), plot.title.position = "plot", legend.key.size = unit(20,"pt"),
                          legend.text= element_text(size=16),
                          axis.title.y = element_text(size=20), axis.title.x = element_text(size=20),
                          axis.text.x = element_text(size=16), axis.text.y = element_text(size=16), legend.position = "top")
```